---
title: "Programming Background"
author: "Evan Brown"
date: "6/7/2022"
---

```{r, echo = FALSE, eval = TRUE}
knitr::opts_chunk$set(fig.path = "../images/")
```

``` {r, echo = FALSE, eval = FALSE}
rmarkdown::render("C:/Users/Ebrown08/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/ST558_Git/_Rmd/2022-06-07-evanbrown-second-blog-post.Rmd", 
                  output_format = "github_document", 
                  output_dir = "../_posts/",
                  output_options = list(
                    keep_html = FALSE
                    )
)
```


## My Programming Background

*Everyone in this course had some programming experience coming in (due to the prerequisites)...*
*What are your thoughts on R vs whatever other software you've used?*
*What functionality do you like about R?*
*What parts do you miss about your other language?*
*Do you consider R a difficult language to learn? (If you knew R prior to the course, describe your experience when first learning it.)*


- My only programming experience prior to ST558 (R) is in ST555 (SAS). Initially, I thought R was much more complex and certainly didn't prefer it. However, I've come to see to flexibility and open-ended design of R that makes it much less rigid than SAS. 

- Personally, my favorite functionality of R so far is markdown. I am impressed with the reproducibility of making documents and how structured output can be with relatively minimal input. I have found public markdown templates for resumes, manuscripts, etc that are useful in the science sector and interesting to play around with. I also like that I can create my own templates and update as needed to fit my goals at the time. Additionally, I like the attitude surrounding R. SAS felt much heavier on the specifics since there was usually just one, maybe two ways to do something. With multiple ways to do things in R, it feels much lighter and allows me to focus on the output. There was a load of SAS help from SAS, but there has also been a ton of R help online as well.

- I miss the organization of SAS code. Writing code in steps felt much more structured than the line by line action R can take on. I also felt like errors had better context in SAS than they do in R, but that could be a lack of experience on my part.

- Learning object oriented programming after learning SAS can be tricky at first. Sometimes it's hard to conceptualize entire datasets as objects and not just row by row, input to output reading. Getting used to packages can be tricky at first also. Since SAS is all inclusive, it was a challenge having to download 15 packages with different functions or syntax for each. However, I do like that we're able to stick with the tidyverse as a way to keep consistent. At work, data.table and ggplot are the two packages we primarily use and once familiarizing myself with those and doing some practical applications with example code, it all started to click. I have picked up the language and Rstudio much quicker than I picked up SAS for sure, probably thanks to my experience with SAS. 



## Example R Markdown Output

Here is a plot of the iris dataset.
```{r, eval=TRUE, echo=FALSE}
plot(iris)
```

